cvx-flats

### 2020-04-08 neurips-20/cvx-flats.tex
Adding in intuition for some proofs and reorg

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\simplex}{\Delta_\Y}
\newcommand{\prop}[1]{\mathrm{prop}[#1]}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
\newcommand{\ccdim}{\mathrm{ccdim}}

@@

%discrete land
Let $\R$ be a (typically finite) report set and $\Y$ a finite outcome set for now.
Note that we do not necessarily have $\Y = \R$.  
For example, in ranking problems, $\R$ may be all $n!$ permutations over the $n$ outcomes forming $\Y$.
A loss $\ell : \R \to \reals_+^\Y$ is a \emph{discrete loss} if $|\R| < \infty$, and in general, we use $\ell$ to denote our \emph{original loss} for which we want to construct a consistent surrogate.
We denote $\ell(r)_y$ as the $y^{th}$ element of the vector $\ell(r)$, and consider this to be the punishment for reporting $r$ if outcome $y$ is realized.
For a distribution $p \in \simplex$, we denote the expected discrete loss for report $r$ to be $\E_{Y \sim p} \ell(r, Y) := \ell(r; p)$.
When considering surrogate losses, we denote the loss by $L : \reals^d \to \reals_+^\Y$, and typically denote a surrogate report with $u$.

@@

\begin{definition}[Consistent]
	Suppose $f_m : \X \to \R$ is the hypothesis function learned by minimizing empirical training loss over $m$ labeled examples.
	For a distribution $D$ over input and label spaces $\X \times\Y$, we say $\{f_m\}$ is \emph{consistent} with respect to a loss $L: \R \to \reals^\Y_+$ if
	\begin{align*}
	\E_D L(f_m(X), Y) &\to \inf_f \E_D L(f(X), Y)~.~
	\end{align*}
	Moreover, we say a loss and link $(L,\psi)$ are consistent with respect to a discrete loss $\ell$ if 
	\begin{align*}
		\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell(\psi \circ f_m(X), Y) \to \inf_f \E_D \ell(\psi \circ f(X), Y)~.~
	\end{align*}
\end{definition}

@@

\begin{definition}[Property, elicits, level set]
	\jessiet{Don't like the notation here... fix later.}
	A \emph{property} is a set-valued function $\Gamma : \simplex \toto \R$ mapping distributions to reports.
	A loss $L : \R \to \reals_+^\Y$ \emph{elicits} the property $\Gamma$ if,
	\begin{equation}
	\forall p \in \simplex, \;\; \Gamma(p) = \argmin_{r \in \R}\inprod{p}{L(r)}
	\end{equation}
	Moreover, we call a \emph{level set} $\Gamma_r := \{p \in \simplex : r \in \Gamma(p)\}$ to be the set of distributions for which reporting $u$ minimizes the expected loss of the loss eliciting $\Gamma$.
\end{definition}
Note, moreover, that we call a property $\gamma: \simplex \toto \R$ \emph{finite} if $|\R| < \infty$.
Without much loss of generality, we assume that finite properties are \emph{non-redundant}, meaning that for each $\gamma_r$, there is no $\gamma_{r'}$ such that $\gamma_{r'} \subseteq \gamma_r$.

In particular, when $\R \subseteq \reals^k$ for some $k \in \mathbb{Z}$, \cite{frongillo2015elicitation} introduce the notion of \emph{Elicitation Complexity} and \emph{Identification Complexity}, which uses $k$ to measure the ``efficiency'' of a loss $L$ eliciting $\Gamma$.
However, their notion of complexity relies not just on the property $\Gamma$, but instead on any property \emph{indirectly elicited} by $L$.

@@

In Proposition~\ref{prop:consistent-implies-indir-elic} below, we relate consistency of a surrogate and link to indirect elicitation.

\section{Implications of property elicitation for consistency}
\begin{proposition}\label{prop:consistent-implies-indir-elic}
	If a surrogate and link pair$(L, \psi)$ is consistent with respect to $\ell$, then $L$ indirectly elicits $\gamma := \prop{\ell}$.
\end{proposition}

\begin{proposition}
	Consider the discrete loss $\ell : \R \to \reals^\Y_+$ and $\gamma:= \prop{\ell}$.
	Then $\eliccvx(\gamma) \leq \ccdim(\ell)$.
	\jessie{Move this to section 5?}
	\jessie{Need to define $\eliccvx$ and $\ccdim$.}
\end{proposition}
\begin{proof}
	Let $d$ be the convex calibration dimension of $\ell$, and $(L, \psi)$ be calibrated with respect to $\ell$ with $L$ convex.
	Then $(L, \psi)$ is also consistent with respect to $\ell$, and therefore $L$ indirectly elicits $\gamma$.
\end{proof}

\section{Characteristics of convex surrogates}\label{sec:char-convex}
\iffalse
\begin{definition}[Identifiable]
	A property $\Gamma: \simplex \to \R$ is \emph{identifiable} if there is a function $V: \R \to \reals^\Y$ such that 
	\begin{equation*}
	\forall p \in \simplex, \; r \in \Gamma(p) \iff V(r) = \vec 0~.~
	\end{equation*}
\end{definition}
\fi

If a property is vector-valued (i.e.$\, \Gamma: \simplex \to \reals^d$), then we can consider the notion of \emph{identifiability}.

\begin{definition}[$d$-identifiable]
	A vector-valued property $\Gamma: \simplex \to \reals^d$ is $d$-identifiable if it holds that $p \in \Gamma_u \iff V(u, p) = \vec 0 \in \reals^d$.
\end{definition}

Moreover, we define the \emph{identification complexity} of a property $\Gamma$ as the minimum dimension $d$ such that there is a link $phi$ such that $\Gamma = \phi \circ \hat \Gamma$ and $\hat \Gamma$ is $d$-identifiable.

@@

When one restricts their focus to \emph{convex} surrogates, we know that all minima are global, and we then have $\nabla L(u, y) = \vec 0$ if and only if $u$ minimizes $L$ in its first component.

\begin{theorem}
	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \to \reals^\Y_+$ such that the infimum $L(\cdot)_y$ is attained for all $y \in \Y$.
	Then for all $p \in \simplex$, we have $u \in \Gamma(p) \iff \vec 0 \in \nabla_{u'} \inprod{p}{L(u')}$.
\end{theorem}
\begin{proof}[Proof sketch]
	Since $L$ is convex, we know that all minimizers are global. 
	This implies that $\vec 0 \in \nabla_{u} L(u; p) \iff u \in \arginf_{u'} L(u';p)$, and the latter is true if and only if $u \in \Gamma(p)$ by definition of elicitation.
\end{proof}

\begin{corollary}
	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \to \reals^\Y_+$ such that the infimum $L(\cdot)_y$ is attained for all $y \in \Y$.
	Then $\Gamma$ is $d$-identifiable.
\end{corollary}

Now we can consider specific characteristics of convex losses in order to understand the convex elicitation complexity \jessie{define?} of a given property.

\begin{theorem}
	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma:= \prop{\ell}$.
	Fix $p \in \simplex$, and let $r$ be the report such that $p \in \gamma_r$.  
	If there is not $(n - d - 1)$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no surrogate loss $L : \reals^d \to \reals^\Y_+$ and link $\psi:\reals^d \to \R$ such that $(L, \psi)$ is calibrated with respect to $\ell$.
	Moreover, this must hold for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{theorem}
\begin{proof}[Proof intuition]
	\jessie{TODO}
\end{proof}

\section{Finite outcome settings and finite properties}
\begin{definition}[Calibrated]
	Let $\ell : \R \to \reals^\Y_+$ be a discrete loss eliciting the property $\gamma$.
	A surrogate loss $L : \reals^d \to \reals^\Y_+$ is \emph{calibrated} with respect to $\ell$ if there exists a link function $\psi: \reals^d \to \R$ such that
	\begin{equation}
	\forall p \in \simplex: \inf_{u \in \reals^d : \psi(u) \not \in \gamma(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.~
	\end{equation}
\end{definition}

\cite{ramaswamy2016convex} introduces \emph{Convex Calibration Dimension}, which studies the minimal dimension input $d$ to construct a convex surrogate $L : \reals^d \to \reals^\Y_+$ that is calibrated with respect to a given discrete loss $\ell$.

\begin{theorem}[\cite{ramaswamy2016convex}]
	If $\Y$ is finite, then the surrogate loss $L:\reals^d \to \reals^\Y_+$ is calibrated with respect to $\ell: \R \to \reals^\Y_+$ if and only if there exists a link function $\psi : \reals^d \to \R$ such that for all distributions $D$ on $\X \times\Y$ and all sequences of (vector) functions $f_m : \X \to \reals^d$,
	\begin{equation*}
	\E_D L(f_m(X), Y) \to \inf_f \E_P L(f(X), Y) \implies \E_D \ell(\psi  \circ f_m(X), Y) \to \inf_f \E_D \ell(\psi \circ f(X), Y)~.~
	\end{equation*}
\end{theorem}
In words, we have that a surrogate and link are calibrated with respect to a discrete loss if and only if any consistent sequence of hypotheses for the surrogate, when linked, is consistent for the discrete loss.

@@

>>>>

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\simplex}{\Delta_\Y}
\newcommand{\prop}[1]{\mathrm{prop}[#1]}
\newcommand{\elic}{\mathrm{elic}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
\newcommand{\ccdim}{\mathrm{ccdim}}

@@

%discrete land
Let $\R$ be a (typically finite) report set and $\Y$ a finite outcome set for now.
Note that we do not necessarily have $\Y = \R$.  
For example, in ranking problems, $\R$ may be all $n!$ permutations over the $n$ outcomes forming $\Y$.
A loss $\ell : \R \times \Y \to \reals_+$ is a \emph{discrete loss} if $|\R| < \infty$, and in general, we use $\ell$ to denote our \emph{original loss} for which we want to construct a consistent surrogate.
We denote $\ell(r)_y$ as the $y^{th}$ element of the vector $\ell(r)$, and consider this to be the punishment for reporting $r$ if outcome $y$ is realized.
For a distribution $p \in \simplex$, we denote the expected discrete loss for report $r$ to be $\E_{Y \sim p} \ell(r, Y) := \ell(r; p)$.
When considering surrogate losses, we denote the loss by $L : \reals^d \times \Y \to \reals_+$, and typically denote a surrogate report with $u$.

@@

\begin{definition}[Consistent]
	Suppose $f_m : \X \to \R$ is the hypothesis function learned by minimizing empirical training loss over $m$ labeled examples.
	For a distribution $D$ over input and label spaces $\X \times\Y$, we say $\{f_m\}$ is \emph{consistent} with respect to a loss $L: \R \times \Y \to \reals_+$ if
	\begin{align*}
	\E_D L(f_m(X), Y) &\to \inf_f \E_D L(f(X), Y)~.~
	\end{align*}
	Moreover, we say a loss and link $(L,\psi)$ are consistent with respect to a discrete loss $\ell$ if 
	\begin{align*}
		\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell(\psi \circ f_m(X), Y) \to \inf_f \E_D \ell(\psi \circ f(X), Y)~.~
	\end{align*}
\end{definition}

@@

\begin{definition}[Property, elicits, level set]
	\jessiet{Don't like the notation here... fix later.}
	A \emph{property} is a set-valued function $\Gamma : \simplex \toto \R$ mapping distributions to reports.
	A loss $L : \R \times \Y \to \reals_+$ \emph{elicits} the property $\Gamma$ if,
	\begin{equation}
	\forall p \in \simplex, \;\; \Gamma(p) = \argmin_{r \in \R}\inprod{p}{L(r)}
	\end{equation}
	Moreover, we call a \emph{level set} $\Gamma_r := \{p \in \simplex : r \in \Gamma(p)\}$ to be the set of distributions for which reporting $u$ minimizes the expected loss of the loss eliciting $\Gamma$.
\end{definition}
Note, moreover, that we call a property $\gamma: \simplex \toto \R$ \emph{finite} if $|\R| < \infty$.
Without much loss of generality, we assume that finite properties are \emph{non-redundant}, meaning that for each $\gamma_r$, there is no $\gamma_{r'}$ such that $\gamma_{r'} \subseteq \gamma_r$.

In particular, when $\R \subseteq \reals^d$ for some $d \in \mathbb{Z}$, \cite{frongillo2015elicitation} introduce the notion of \emph{Elicitation Complexity} and \emph{Identification Complexity}, which uses $d$ to measure the ``efficiency'' of a loss $L$ eliciting $\Gamma$.
However, their notion of complexity relies not just on the property $\Gamma$, but instead on any property \emph{indirectly elicited} by $L$.

@@

In Proposition~\ref{prop:consistent-implies-indir-elic}, we relate indirect elicitation to the consistency of a surrogate loss.

\subsubsection{Elicitation Complexity}
When properties are vector-valued (i.e.$\, \Gamma: \simplex \to \reals^d$),~\cite{frongillo2015elicitation} introduce the notion of \emph{elicitation complexity}.

\begin{definition}[$d$-elicitable, Elicitation Complexity]
	We say a elicitable property $\Gamma:\simplex \to \reals^d$ is $d$-elicitable if there is a loss $L : \reals^d \times \Y \to \reals_+$ eliciting it.
	The \emph{elicitation complexity} of a property $\Gamma := \elic(\Gamma)$ is the minimum dimension $d$ such that there is a property $\hat \Gamma$ that is $d$-elicitable and link $\psi$ so that $\Gamma = \psi \circ \hat \Gamma$.
	Moreover, we use $\eliccvx(\Gamma)$ to denote the elicitation complexity of $\Gamma$ when the directly elicited property $\hat \Gamma$ is elicited by a \emph{convex} surrogate loss.
\end{definition}

\iffalse
\begin{definition}[Identifiable]
	A property $\Gamma: \simplex \to \R$ is \emph{identifiable} if there is a function $V: \R \times\Y \to \reals$ such that 
	\begin{equation*}
	\forall p \in \simplex, \; r \in \Gamma(p) \iff V(r) = \vec 0~.~
	\end{equation*}
\end{definition}
\fi

Similarly, we can consider the notion of \emph{identifiability} for vector-valued properties.

\begin{definition}[$d$-identifiable, Identification Complexity]
	A vector-valued property $\Gamma: \simplex \to \reals^d$ is $d$-identifiable if it holds that $p \in \Gamma_u \iff V(u, p) = \vec 0 \in \reals^d$.
	Like elicitation complexity, we define the \emph{identification complexity} of a property $\Gamma$ as the minimum dimension $d$ such that there is a link $\psi$ such that $\Gamma = \psi \circ \hat \Gamma$ and $\hat \Gamma$ is $d$-identifiable.
\end{definition}

@@

\section{Implications of property elicitation for consistency}
\begin{proposition}\label{prop:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to $\ell$, then $L$ indirectly elicits $\gamma := \prop{\ell}$.
\end{proposition}
\begin{proof}
Since consistency must hold for all $D$ over $\X \times \Y$, let us focus on distributions $D$ such that the probability of any $x \in \X$ is a point mass, allowing us to focus on conditional distributions $p \in \simplex$ such that $p = Pr[Y = y | X = x]$, allowing us to be agnostic to $\X$.
Now, instead of focusing on sequences of optimal hypotheses, we can focus on optimal reports when given $x \in \X$ if the hypothesis class $\F$ is sufficiently rich (i.e. the optimal hypothesis is in the hypothesis class for all $x \in \X$.)
 \jessie{FINISH}
\end{proof}

This result allows us to apply the upper bounds from~\cite{frongillo2015elicitation} when trying to find the minimal dimension consistent surrogate for a given loss.
In particular, when $\Y$ is finite and we restrict $L$ to be convex, this translates to applications to \emph{convex calibration dimension} of~\cite{ramaswamy2016convex}.

\section{Characteristics of convex surrogates}\label{sec:char-convex}
When one restricts their focus to \emph{convex} surrogates, we know that all minima are global, and we then have $\nabla L(u, y) = \vec 0$ if and only if $u$ minimizes $L$ in its first component.

\begin{theorem}
	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \times \Y \to \reals_+$ such that the infimum $L(\cdot)_y$ is attained for all $y \in \Y$.
	Then for all $p \in \simplex$, we have $u \in \Gamma(p) \iff \vec 0 \in \nabla_{u'} \inprod{p}{L(u')}$.
\end{theorem}
\begin{proof}[Proof sketch]
	Since $L$ is convex, we know that all minimizers are global. 
	This implies that $\vec 0 \in \nabla_{u} L(u; p) \iff u \in \arginf_{u'} L(u';p)$, and the latter is true if and only if $u \in \Gamma(p)$ by definition of elicitation.
\end{proof}

In a large sense, this lets us conclude that $d$-convex elicitable properties are also $d$-identifiable, taking the identification function $V(\cdot;p)$ to be $\nabla_u L(\cdot; p)$ for all $p \in \simplex$.
\begin{corollary}
	Suppose the property $\Gamma$ is elicited by the convex loss $L : \reals^d \times \Y \to \reals_+$ such that the infimum $L(\cdot, y)$ is attained for all $y \in \Y$.
	Then $\Gamma$ is $d$-identifiable.
\end{corollary}

Now we can consider specific characteristics of convex losses in order to understand the convex elicitation complexity of a given property.

\begin{theorem}
	Suppose we are given a discrete loss $\ell:\R \times \Y \to \reals_+$ eliciting $\gamma:= \prop{\ell}$.
	Fix $p \in \simplex$, and let $r$ be the report such that $p \in \gamma_r$.  
	If there is not $(n - d - 1)$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no surrogate loss $L : \reals^d \times \Y \to \reals_+$ and link $\psi:\reals^d \to \R$ such that $(L, \psi)$ is calibrated with respect to $\ell$.
	Moreover, this must hold for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{theorem}
\begin{proof}[Proof intuition]
	\jessie{TODO}
\end{proof}

\section{Finite outcome settings and finite properties}
\begin{definition}[Calibrated]
	Let $\ell : \R \times \Y \to \reals_+$ be a discrete loss eliciting the property $\gamma$.
	A surrogate loss $L : \reals^d \times \Y \to \reals_+$ is \emph{calibrated} with respect to $\ell$ if there exists a link function $\psi: \reals^d \to \R$ such that
	\begin{equation}
	\forall p \in \simplex: \inf_{u \in \reals^d : \psi(u) \not \in \gamma(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.~
	\end{equation}
\end{definition}

\cite{ramaswamy2016convex} introduces \emph{Convex Calibration Dimension}, which studies the minimal dimension input $d$ to construct a convex surrogate $L : \reals^d \times \Y \to \reals_+$ that is calibrated with respect to a given discrete loss $\ell$.

\begin{theorem}[\cite{ramaswamy2016convex}]\label{thm:calib-iff-consistent}
	If $\Y$ is finite, then the surrogate loss $L:\reals^d \to \reals^\Y_+$ is calibrated with respect to $\ell: \R \times \Y \to \reals_+$ if and only if there exists a link function $\psi : \reals^d \to \R$ such that for all distributions $D$ on $\X \times\Y$ and all sequences of (vector) functions $f_m : \X \to \reals^d$,
	\begin{equation*}
	\E_D L(f_m(X), Y) \to \inf_f \E_P L(f(X), Y) \implies \E_D \ell(\psi  \circ f_m(X), Y) \to \inf_f \E_D \ell(\psi \circ f(X), Y)~.~
	\end{equation*}
\end{theorem}
In words, we have that a surrogate and link are calibrated with respect to a discrete loss if and only if any consistent sequence of hypotheses for the surrogate, when linked, is consistent for the discrete loss.

@@

Theorem~\ref{thm:calib-iff-consistent} allows us to now relate convex elicitation complexity and convex calibration dimension introduced by~\cite[Definition 10]{ramaswamy2016convex}.
\begin{proposition}
	Consider the discrete loss $\ell : \R \times \Y \to \reals_+$ and $\gamma:= \prop{\ell}$.
	Then $\eliccvx(\gamma) \leq \ccdim(\ell)$.
	\jessie{Move this to section 5?}
	\jessie{Need to define $\eliccvx$ and $\ccdim$.}
\end{proposition}
\begin{proof}
	Let $d$ be the convex calibration dimension of $\ell$, and $(L, \psi)$ be calibrated with respect to $\ell$ with $L$ convex.
	Then $(L, \psi)$ is also consistent with respect to $\ell$, and therefore $L$ indirectly elicits $\gamma$.
\end{proof}
